---
title: "Weather and Corn Yield Regressions"
author: "Nick Gubbins"
date: "2/25/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
```

# Corn and Weather Data Analysis

## Loading in data

### Load the PRISM daily maximum temperatures

```{r tmax data}

# daily max temperature
# dimensions: counties x days x years
prism <- readMat("data/prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]
plot(1:366, t_1981_c1, type = "l")

ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (°C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")


```
```{r tidying up}

# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years)

# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf)

```

## Investigating temperature trends

### Summer temperature trends: Winneshiek County

```{r temp trends}

tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

winnesummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnesummer, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax)

```

### Winter Temperatures - Winneshiek County

```{r winter temps}

winnewinter <- tmaxdf %>%
  filter(countyfp==191 & doy <= 59 | doy >= 335 & !is.na(tmax)) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnewinter, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax)

```

### Multiple regression -- Quadratic time trend

```{r quadratic temp trend}

winnewinter$yearsq <- winnewinter$year^2

lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
summary(lm_wintertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values

ggplot(winnewinter) +
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "tmax")

```

### Download NASS corn yield data

```{r yield download, include=F, message=F}

# set our API key with NASS
nassqs_auth(key = "6391961E-A70F-3C2C-B3AD-97510C837921")

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

```

## Assignment

### Q1 

#### Part A

Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend

```{r}
#glimpse(cornyields)
library(lubridate)
data <- cornyields %>%
  filter(county_name == "WINNESHIEK")

m1a <- lm(data = data, yield~year)
summary(m1a)

data %>%
  ggplot(., aes(x = year, y = yield))+
  geom_point()+
  geom_line()+
  labs(x = 'Year', y = 'Corn Yield (T)', title= 'Corn Yields in Winnesiek County')

```

Yes, there is a significant, postitive trend over time.

#### Part B

Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 

```{r}
data$year_sq <- data$year^2
m1b <- lm(data = data, yield~year+year_sq)
summary(m1b)

plot(m1b, which = 1)
```

Yes, there is evidence that the trend is slowing in more recent years.

### Q2

Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.

```{r}
data_t <- data %>%
  left_join(., tmaxdf, by = 'year') %>%
  filter(is.finite(tmax))

m2a <- lm(data = data_t, yield~tmax)
summary(m2a)

#plot(m2a)

data_t$tmax_sq <- data_t$tmax^2

m2b <- lm(data = data_t, yield~tmax+tmax_sq)
summary(m2b)

#plot(m2b)

p <- ggplot(data_t,aes(x=yield, y = tmax)) +
  geom_bin2d(binwidth = 1)
p

p <- ggplot(data_t,aes(x=yield, y = tmax_sq)) +
  geom_bin2d(binwidth = 1)
p
```

Yes, adding the square of max temperature improved our model somewhat. 

### Q3

Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.

```{r}
data3 <- cornyields %>%
  mutate(countyfp = as.factor(county_ansi)) %>%
  full_join(., tmaxdf, by = c('year', 'countyfp')) %>%
  mutate(tmax_sq = tmax^2,
         countyfp = as.factor(countyfp))

summary(lm(data = data3, yield~0+tmax+tmax_sq))
  
```

Yes,there is a positive relationship between max temperature and growth for the year, as indicated by the low p-value on our slope. However, max temperature only explains some of the variability in our data, as indicated by an r-squared value of 0.54.

### Q4

Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.

```{r}
m4 <- lm(data = data3, yield~0+tmax+tmax_sq*countyfp)
summary(m4)

# Plot is breaking my html, but this code generates it.
#plot(m4, which = 1)

```

The model with county code as a factor still has the temperature terms as highly significant, just with differences in degree between counties.

### Q5

Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.


```{r}
# parameters to query on 
params <- list(commodity_desc = "SOYBEANS", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
soyyieldsall <- nassqs_yields(params)

soyyieldsall$county_ansi <- as.numeric(soyyieldsall$county_ansi)
soyyieldsall$yield <- as.numeric(soyyieldsall$Value)

# clean and filter this dataset
soyyields <- select(soyyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
soyyields <- tibble(soyyields)

# reduce to buena vista county
buena_soy <- soyyields %>%
  filter(county_name == 'BUENA VISTA')

ggplot(buena_soy, aes(x = year, y = yield))+
  geom_point()+
  geom_line()+
  labs(x = 'Year', y = 'Corn Yield (T)', title= 'Corn Yields in Buena Vista County')

m5 <- lm(data = buena_soy, yield~year)
  
```


